---
title: "R Notebook"
output: html_notebook

---

setwd("~/career_da-pffs/udacity-eda/eda-course-materials/lesson6")


```{r}
install.packages(ggplot2)
install.packages("reshape")
install.packages("tidyr")
install.packages("dplyr")
library(dplyr)
library(tidyr)
library(ggplot2)

```


```{r}
gdURL <- "http://www.stat.ubc.ca/~jenny/notOcto/STAT545A/examples/gapminder/data/gapminderDataFiveYear.txt"
gDat <- read.delim(file = gdURL)
gDat <- droplevels(subset(gDat, continent != "Oceania"))
library(ggplot2)
library(plyr)
# http://rpubs.com/neil_spencer/stat545a-2013-hw05_spencer-nei
```


```{r}
str(gDat)
```

```{r}
#summary(gDat)

cor.test(gDat$lifeExp, gDat$gdpPercap, method = 'pearson') 
```
Observation - a strong positive correlation.


summary(gDat$lifeExp)
  Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
  23.60   48.08   60.34   59.26   70.75   82.60 



1. A scatter plot of GDPPercap vs LifeExp
```{r}
ggplot(aes(x = lifeExp, y = gdpPercap, color = continent), data = gDat) + 
  geom_point() +
  ggtitle("GdpPercap vs. LifeExp") +
  labs(x="lifeExp", y="gdpPercap") +
  scale_x_continuous(breaks = seq(20,90,4)) 
```
Observation - A few outliers - a country with almost 0 gdpPercap and some with extremely high GDP
Most countries with a gdp below 15000.  To reduce overploting with alpha and coordinate transormation 

min(gDat$gdpPercap)

```{r}
ggplot(aes(x = lifeExp, y = gdpPercap, color = continent), data = gDat) + 
  geom_point(alpha = 1/2) +
  ggtitle("GdpPercap vs. LifeExp") +
  labs(x="lifeExp", y="gdpPercap") +
  scale_x_continuous(breaks = seq(20,90,4)) + 
  coord_trans(x="sqrt", y='sqrt')
```
# Omiting the top 1% of values by applying the quartile method


```{r - Omiting the top 1%}
ggplot(aes(x = lifeExp, y = gdpPercap, color = continent),
       data = gDat) + 
  geom_point(alpha = 1/2) +
  ggtitle("GdpPercap vs. LifeExp") +
  labs(x="lifeExp", y="gdpPercap") +
  scale_x_continuous(breaks = seq(20,90,4)) + 
  coord_trans(x="sqrt", y='sqrt') + 
   xlim(20, quantile(gDat$lifeExp, 0.99)) +
  ylim(0, quantile(gDat$gdpPercap, 0.99)) +
  geom_smooth(method="lm", color="red")
```
Geom-smooth omits most of the values therefore does not seem the proper choice for this data


Making the plot more readable by creating and using conditoning summaries.  
Error below - I do not understand it??

```{r}
gDat_by_country<- gDat %>% 
                group_by(country) %>%
                summarise(lifeExp_mean=mean(lifeExp),
                lifeExp_median=median(lifeExp),
                 n=n()) %>%
                arrange(country)
          head(gDat_by_country)
```



```{r Plotting Conditional Summaries}

ggplot(aes(x = continent, y = lifeExp),
       data = subset(gDat, !is.na(continent))) + 
  geom_boxplot() +
  stat_summary(geom = 'point', shape = 4, fun.y = mean)

```

```{r}
ggplot(aes(x = continent, y = gdpPercap, color = continent),
       data = subset(gDat, !is.na(continent))) + 
  geom_boxplot() +
  stat_summary(geom = 'point', shape = 4, fun.y = median) +
  scale_y_continuous(limits = c(0, 30000),breaks=seq(0, 30000, 1000))
  
```



Write code to create a new data frame - why??

```{r}

year_lifeExp_group <- group_by(gDat, year, lifeExp) %>% 
                filter(!is.na(year)) %>% 
                summarise(mean_gdpPercap = mean(gdpPercap),
                          median_gdpPercap = median(gdpPercap),
                          n = n())
ungroup() %>%
arrange(gDat_year_lifeExp_group, year)
```
```
why?










