---
output: html_document
editor_options: 
  chunk_output_type: inline
---
clive kmg
18/10/2018

Lesson 5
========================================================
```{r}
setwd("~/career_da-pffs/udacity-eda/eda-course-materials/lesson5")
```


```{r Third Qualitative Variable}
pf <- read.csv("pseudo_facebook.tsv", sep = '\t')
```

***
```{r}

library(readr)
library(ggplot2)
library(dplyr)
library(tidyr)
library(data.table)
```

### Multivariate Data
Notes: This lesson goes beyond bivariate analysis.

Moira Perceived Audience Size Colored by Age
Notes: Adding color to represent age, but there was no pattern. Too much overplotting, which makes color as an encoder less effective.

#Third Qualitative Variable
Notes: Conducting EDA can lead to deadends. Don't panic.

***

### Moira Perceived Audience Size Colored by Age
Notes:

```{r Plotting Conditional Summaries}

ggplot(aes(x = gender, y = age),
       data = subset(pf, !is.na(gender))) + 
  geom_boxplot() +
  stat_summary(geom = 'point', shape = 4, fun.y = mean)

ggplot(aes(x = age, y = friend_count, color = gender), 
       data = subset(pf, !is.na(gender))) +
  geom_line(aes(color = gender), stat = 'summary', fun.y = median)

```
```
### Third Qualitative Variable
Notes:
# Write code to create a new data frame,
# called 'pf.fc_by_age_gender', that contains
# information on each age AND gender group.

# The data frame should contain the following variables:

#    mean_friend_count,
#    median_friend_count,
#    n (the number of users in each age and gender grouping)

# Here is an example of the structure of your data frame. Your
# data values will be different. Note that if you are grouping by
# more than one variable, you will probably need to call the
# ungroup() function. 

#   age gender mean_friend_count median_friend_count    n
# 1  13 female          247.2953                 150  207
# 2  13   male          184.2342                  61  265
# 3  14 female          329.1938                 245  834
# 4  14   male          157.1204                  88 1201

# See the Instructor Note for two hints.

You can include multiple variables to split the data frame when using group_by() function in the dplyr package.

#new_groupings <- group_by(data, variable1, variable2)

OR
using chained commands...

#new_data_frame <- data_frame %>%
  group_by(variable1, variable2) %>%

Repeated use of summarise() and group_by(): The summarize function will automatically remove one level of grouping (the last group it collapsed).

http://stackoverflow.com/questions/21653295/dplyr-issues-with-group-by
```{r}
library(dplyr)
#chain functions together %>%
pf.fc_by_age_gender <- pf %>%
  filter(!is.na(gender)) %>%  
  group_by(age, gender)  %>%
  summarise(mean_friend_count= mean(friend_count),
            median_friend_count = median(friend_count),
            n = n()) %>%
  ungroup() %>%
  arrange(age)
  
  #use the subset or filter command
```

#or

age_gender_group <- group_by(pf, age, gender)
age_gender_group <- filter(age_gender_group, !is.na(gender))
pf.fc_by_age_gender <- summarise(age_gender_group,
                                 mean_friend_count = mean(friend_count),
                                 median_friend_count = median(friend_count),
                                 n = n())
arrange(pf.fc_by_age_gender, age)
```
```{r}
head(pf.fc_by_age_gender, 3)
```
``
***
### Plotting Conditional Summaries
Notes:
# Create a line graph showing the
# median friend count over the ages
# for each gender. Be sure to use
# the data frame you just created,
# pf.fc_by_age_gender.
Your code should look similar to the code we used to make the plot the first time. It will not need to make use of the stat and fun.y parameters.

```{r Plotting Conditional Summaries}
ggplot(aes(x = age, y = median_friend_count, color = gender), 
       data = pf.fc_by_age_gender) +
  geom_line()
```

***

### Thinking in Ratios
Notes: Looking at the graph especially in the young group how many more times is the average female user more than the average male user-- lets put it in relative terms.  Which leads us to the next topic!
***

### Wide and Long Format
https://s3.amazonaws.com/udacity-hosted-downloads/ud651/DataWranglingWithR.pdf
Notes:

***

### Reshaping Data
https://seananderson.ca/2013/10/19/reshape/
Notes:
Can transform data via dplyr + tidyr or with reshape library.
Reshaping Data
```{r}
install.packages("stringi",type="win.binary") 
library(reshape2)

pf.fc_by_age_gender.wide <- dcast(pf.fc_by_age_gender,
                                  age ~ gender,
                                  value.var = 'mean_friend_count')
head(pf.fc_by_age_gender.wide)
```
```{r}
# same effect with dplyr and tidyr
pf.fc_by_age_gender.wide2 <-
  subset(pf.fc_by_age_gender[c('age', 'gender', 'median_friend_count')],
         !is.na(gender)) %>%
  spread(gender, median_friend_count) %>%
  mutate(ratio = female / male)
```

```{r}
# try out melt
pf.fc_by_age_gender.long <- melt(pf.fc_by_age_gender.wide,
                                 measure.vars = c('male', 'female'),
                                 value.name = 'age')
```

***

### Ratio Plot
Notes:# Plot the ratio of the female to male median
# friend counts using the data frame
# pf.fc_by_age_gender.wide.

# Think about what geom you should use.
# Add a horizontal line to the plot with
# a y intercept of 1, which will be the
# base line. Look up the documentation
# for geom_hline to do that. Use the parameter
# linetype in geom_hline to make the
# line dashed.

# The linetype parameter can take the values 0-6:
# 0 = blank, 1 = solid, 2 = dashed
# 3 = dotted, 4 = dotdash, 5 = longdash
# 6 = twodash

```{r Ratio Plot}
ggplot(aes(x = age, y = female / male),
       data = pf.fc_by_age_gender.wide) + 
  geom_line() + 
  geom_hline(yintercept = 1, alpha = 0.3, linetype = 2)

#or

ggplot(aes(x = age, y = ratio), data = pf.fc_by_age_gender.wide2) +
  geom_line() +
  geom_hline(yintercept = 1, alpha = 0.3, linetype = 'dashed')

```

***

### Third Quantitative Variable
A common mistake is to use tenure rather than pf$tenure or with(pf, tenure...). Remember that you need to access the variable in the data frame.
Notes:# Create a variable called year_joined
# in the pf data frame using the variable
# tenure and 2014 as the reference year.

# The variable year joined should contain the year
# that a user joined facebook.

#Hint 1:
Divide the tenure variable by a number. Tenure is measured in days, but we want to convert it to years. 
#Hint 2: 
Subtract tenure measured in years from 2014. What does the decimal portion represent? Should we round up or round down to the closest year? 
#Hint 3: 
You can use the floor() function to round down to the nearest integer. You can use the ceiling() function to round up to the nearest integer. Which one should you use?

```{r Third Quantitative Variable}
pf$year_joined <- floor(2014 - pf$tenure/365)

length(pf$year_joined)
```

***

### Cut a Variable
The cut function is useful for turning continuous variables into factors. eg
> c1 <- cut(clinical.trial$age, breaks = 4)
```{r}
## specify break points explicitly using seq function
 
## look what seq does  
seq(30, 80, by = 10)   
# ans= 30 40 50 60 70 80 
## cut the age variable using the seq defined above
c1 <- cut(clinical.trial$age, breaks = seq(30, 80, by = 10))
## table of the resulting factor           
table(c1)
```

Notes:
# Create a new variable in the data frame
# called year_joined.bucket by using
# the cut function on the variable year_joined.

# You need to create the following buckets for the
# new variable, year_joined.bucket

#        (2004, 2009]
#        (2009, 2011]
#        (2011, 2012]
#        (2012, 2014]

# Note that a parenthesis means exclude the year and a
# bracket means include the year.

# Look up the documentation for cut or try the link
# in the Instructor Notes to accomplish this task.
http://www.r-bloggers.com/r-function-of-the-day-cut-2/

```{r}
summary(pf$year_joined)
```
```{r}
table(pf$year_joined)
```
= 2005  2006  2007  2008  2009  2010  2011  2012  2013  2014 
    9    15   581  1507  4557  5448  9860 33366 43588    70

***
```{r}

pf$year_joined.bucket <- cut(pf$year_joined, c(2004, 2009, 2011, 2012, 2014))
```


### Plotting it All Together
Notes:

```{r Plotting it All Together}
table(pf$year_joined.bucket, useNA = 'ifany')
```
# = (2004,2009] (2009,2011] (2011,2012] (2012,2014]        <NA> 
       6669       15308       33366       43658           2 
***
```{r}
ggplot(data = subset(pf, !is.na(year_joined.bucket)), 
aes(x= age, y=friend_count)) +
  geom_line(aes(color = year_joined.bucket),
  stat = 'summary', fun.y = median)
```


### Plot the Grand Mean
Notes: Use the grand mean for high-level observation on data.

# Write code to do the following:

# (1) Add another geom_line to code below
# to plot the grand mean of the friend count vs age.

# (2) Exclude any users whose year_joined.bucket is NA.

# (3) Use a different line type for the grand mean.

# As a reminder, the parameter linetype can take the values 0-6:

# 0 = blank, 1 = solid, 2 = dashed
# 3 = dotted, 4 = dotdash, 5 = longdash
# 6 = twodash

```{r Plot the Grand Mean}
ggplot(data = subset(pf, !is.na(year_joined.bucket)),
aes(x= age, y=friend_count)) +
  geom_line(aes(color = year_joined.bucket), 
  stat = 'summary', fun.y = mean) +
  geom_line(stat = 'summary', fun.y = mean, linetype = 2)

```


```{r}

ggplot(aes(x = age, y = friend_count), 
       data = subset(pf, !is.na(pf$year_joined.bucket))) +
  geom_line(aes(color = factor(year_joined.bucket)), stat = 'summary', fun.y = mean) +
  geom_line(color = 'black', linetype = 'dashed', stat = 'summary', fun.y = mean)
```


***

### Friending Rate
Notes:

```{r Friending Rate}
with(subset(pf, tenure >= 1), summary(friend_count / tenure))
```
   Min.  1st Qu.   Median     Mean  3rd Qu.     Max. 
  0.0000   0.0775   0.2205   0.6096   0.5658 417.0000
***

### Friendships Initiated
Notes:

What is the median friend rate? 0.221

What is the maximum friend rate? 417.0

```{r Friendships Initiated
ggplot(aes(x = tenure, y = friendships_initiated /tenure),
       data = subset(pf, tenure >= 1)) +
    geom_line(aes(color = year_joined.bucket),)
    
    
```{r}

```

```

***

### Bias-Variance Tradeoff Revisited
Understanding the Bias-Variance Tradeoff
http://scott.fortmann-roe.com/docs/BiasVariance.html

NOTE: The code changing the binning is substituting x = tenure in the plotting expressions with x = 7 * round(tenure / 7), etc., binning values by the denominator in the round function and then transforming back to the natural scale with the constant in front.


Notes: To get rid of the noise you have to vary the bin size differently eg from change tenure from 7, 30 to 90 and see the difference 

```{r Bias-Variance Tradeoff Revisited}
library(gridExtra)
p1 <- ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

p2 <- ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

p3 <- ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

p4 <- ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)
grid.arrange(p1, p2, p3, p4, ncol=1)
ggsave('biastradeoff.png')

```
```{r}
ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)
```

# Instead of geom_line(), use geom_smooth() to add a smoother to the plot.
# You can use the defaults for geom_smooth() but do color the line
# by year_joined.bucket

```{r}
ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_smooth(aes(color = year_joined.bucket))
```

Friendships initiated decreases as tenure increases!
***

### Sean's NFL Fan Sentiment Study
Notes:Notes: It took averaging to get the story to come out for fan sentiment related to NFL team wins and losses. If data is too noisy, average until you see a reasonable story or a story you'd expect. As he added more data, he had lower variance but more biased data.
interesting!!!
***

### Introducing the Yogurt Data Set
Notes:
#read in the data

```{r}
#rm('yogurt')
#yo <- read.csv('yogurt.csv')

#str(yo)

# change the id from an int to a factor
#yo$id <- factor(yo$id)
str(yo)

```

***

### Histograms Revisited
Notes:If the binwidth is increased, the discreteness of the data disappears, creating a biased model. You can average too much, which would lead to biasness depending on the underlying characteristics of the data set
left long tailed data with bw = 10


```{r Histograms Revisited}
ggplot(aes(x= price), data = yo) +
  geom_histogram(binwidth = 1)  

```
Notice - there is discretness in the plot.  some prices have many observations while adjacent ones have nothing
-with big binwidth like 10 we have no discretness - details are obscurred
There is price variation for the products.
For this data, this histogram is a very biased model since change of binwidth obscures the picture of the data.  Are there other ways we could have noticed this issue....?
Like other exploratory methods
***
```{r}
#shows distinct prices/types
unique(yo$price)
length(unique(yo$price))
table(yo$price)
#gives you a picture simr to the histogram
```


### Number of Purchases
Notes:# Create a new variable called all.purchases,
# which gives the total counts of yogurt for
# each observation or household.

# One way to do this is using the transform
# function. You can look up the function transform
# and run the examples of code at the bottom of the
# documentation to figure out what it does.

# The transform function produces a data frame
# so if you use it then save the result to 'yo'!

# OR you can figure out another way to create the
# variable.

?transform()

```{r Number of Purchases}
yo <- transform(yo, all.purchases = strawberry + blueberry + pina.colada + plain + mixed.berry)

summary(yo$all.purchases)

qplot(x = all.purchases, data = yo, binwidth = 1,
      fill = I('#099DD9'))
```

***

### Prices over Time
Notes:
# Create a scatterplot of price vs time.

# This will be an example of a time series plot.

# Resolve overplotting issues by using
# techniques you learned in Lesson 4.

# What are some things that you notice?

```{r Prices over Time}
ggplot(aes(x = time, y= price), data = yo) +
  geom_jitter(alpha = 1/5, shape = 21)
  
  #or
ggplot(data = yo, aes(x = time, y = price)) + 
  geom_jitter(alpha = 1/4, shape = 21, fill = I('#F79420'))

```
Prices were generally higher at the end of the time period. There was some lumpiness in the data, which could be indicative of possible coupon usage, or experimentation in differently priced yogurts. A few outliers had lower prices towards the end of the time period.
lower prices may be experiments till the right higher price is set
***

### Sampling Observations
Notes: best way to get a good fill with a dataset

***

### Looking at Samples of Households
Note: x %in% y returns a logical (boolean) vector the same length as x that says whether each entry in x appears in y. That is, for each entry in x, it checks to see whether it is in y. This allows us to subset the data so we get all the purchases occasions for the households in the sample. Then, we create scatterplots of price vs. time and facet by the sample id.

Use the pch or shape parameter to specify the symbol when plotting points. Scroll down to 'Plotting Points' on QuickR's Graphical Parameters
https://www.statmethods.net/advgraphs/parameters.html

```{r Looking at Sample of Households}
set.seed(1026)
sample_id <- unique(yo$id)
sample.ids <- sample(x = sample_id, size = 16)

ggplot(aes(x = time, y = price),
       data = subset(yo, id %in% sample.ids)) +
  facet_wrap( ~ id) +
  geom_line() +
  geom_point(aes(size = all.purchases), pch = 1)
```

***

### The Limits of Cross Sectional Data
Notes:

***

### Many Variables
Notes:

***

### Scatterplot Matrix
Notes:

***

### Even More Variables
Notes:

***

### Heat Maps
Notes:

```{r}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

