---
title: "R Notebook"
output: html_notebook
example - https://github.com/GoldbergData/Lesson-8-Problem-Set-Explore-Many-Variables/blob/master/Problem_Set_Explore_Many_Variables.md
---

setwd("~/career_da-pffs/udacity-eda/eda-course-materials/lesson6")

```{r install the packages}

install.packages(ggplot2)
install.packages('reshape')
install.packages('plyr')
install.packages('GGally')
```


```{r}
install.packages('read.xlsx')
library(readxl)
```


```{r load the packages and data}
library(ggplot2)
library(reshape)
library(dplyr)
library(GGally)

data(diamonds)
```


```{r - get familiar with the data}
str(diamonds)
names(diamonds)

```

```{r}
ggcorr(diamonds, palette = "RdBu", label = TRUE)
```


Quiz 1 - Price Histograms with Facet and Color

# Create a histogram of diamond prices.
# Facet the histogram by diamond color
# and use cut to color the histogram bars.

# The plot should look something like this.
# http://i.imgur.com/b5xyrOu.jpg

# Note: In the link, a color palette of type
# 'qual' was used to color the histogram using
# scale_fill_brewer(type = 'qual')

```{r}
ggplot(aes(x = log10(price), fill = cut), data = diamonds) + 
  geom_histogram(binwidth = 0.1) +
  scale_fill_brewer(type = 'qual') + 
   facet_wrap( ~ color)
ggsave('pricehistogram.png')
```
labs(x = 'Price', y = 'Frequency',
       title = 'Price of Diamonds by Color and Cut') +

Quiz 2 - Price vs. Table Colored by Cut
# Create a scatterplot of diamond price vs.
# table and color the points by the cut of
# the diamond.

# The plot should look something like this.
# http://i.imgur.com/rQF9jQr.jpg

# Note: In the link, a color palette of type
# 'qual' was used to color the scatterplot using
# scale_color_brewer(type = 'qual')

```{r}
ggplot(aes(x = table, y = price), data = diamonds) + 
geom_point(aes(color = cut)) + 
   labs(x = 'Price', y = 'Frequency',
       title = 'Price of Diamonds by Color and Cut') +
  scale_color_brewer(type = 'qual') + 
  scale_x_continuous(breaks = seq(50, 80, 2))

#http://i.imgur.com/rQF9jQr.jpg
```
What is the typical table range for the majority of diamonds of ideal cut? 53-57. What is the typical table range for the majority of diamonds of premium cut? 58-62.

Quiz 3 - Price vs. Volume and Diamond Clarity

# Create a scatterplot of diamond price vs.
# volume (x * y * z) and color the points by
# the clarity of diamonds. Use scale on the y-axis
# to take the log10 of price. You should also
# omit the top 1% of diamond volumes from the plot.

# Note: Volume is a very rough approximation of
# a diamond's actual volume.

# The plot should look something like this.
# http://i.imgur.com/excUpea.jpg

# Note: In the link, a color palette of type
# 'div' was used to color the scatterplot using
# scale_color_brewer(type = 'div')

```{r}
volume <- diamonds$x*diamonds$y*diamonds$z

ggplot(data = diamonds, aes(x = volume, y = price)) + 
  geom_point(aes(colour = clarity)) + 
  scale_y_log10() + 
  coord_cartesian(xlim=c(0,quantile(volume, 0.99))) +
  scale_color_brewer(type = 'div')
```

Quiz 5 - Proportion of Friendships Initiated
# Many interesting variables are derived from two or more others.
# For example, we might wonder how much of a person's network on
# a service like Facebook the user actively initiated. Two users
# with the same degree (or number of friends) might be very
# different if one initiated most of those connections on the
# service, while the other initiated very few. So it could be
# useful to consider this proportion of existing friendships that
# the user initiated. This might be a good predictor of how active
# a user is compared with their peers, or other traits, such as
# personality (i.e., is this person an extrovert?).

# Your task is to create a new variable called 'prop_initiated'
# in the Pseudo-Facebook data set. The variable should contain
# the proportion of friendships that the user initiated.

```{r}
library(readr)
pf <- read_tsv('pseudo_facebook.tsv')

```

```{r}
pf$prop_initiated <- pf$friendships_initiated / pf$friend_count
```

Quiz 6 - prop_initiated vs. tenure

# Create a line graph of the median proportion of
# friendships initiated ('prop_initiated') vs.
# tenure and color the line segment by
# year_joined.bucket.

# Recall, we created year_joined.bucket in Lesson 5
# by first creating year_joined from the variable tenure.
# Then, we used the cut function on year_joined to create
# four bins or cohorts of users.

# (2004, 2009]
# (2009, 2011]
# (2011, 2012]
# (2012, 2014]

# The plot should look something like this.
# http://i.imgur.com/vNjPtDh.jpg
# OR this
# http://i.imgur.com/IBN1ufQ.jpg

?floor() = rounding of numbers

NOTE: You may need to filter out missing data from your dataframe before you are able to create your plot.

***
pf <- read.csv('pseudo_facebook.tsv', sep = '\t')
pf$prop_initiated = pf$friendships_initiated / pf$friend_count
pf$prop_initiated[is.nan(pf$prop_initiated)] <- 0    ???


```{r}
pf$year_joined <- floor(2014 - (pf$tenure / 365))

pf$year_joined.bucket <- cut(pf$year_joined, breaks =
                             c(2004, 2009, 2011, 2012, 2014))
```

```{r}
library(ggplot2)

ggplot(aes(x = tenure, y = prop_initiated), data = pf) + 
  labs(x = 'Tenure', 
       y = 'Median Proportion of Friends Initiated',
       title = 'Proportion Initiated vs. Tenure') + geom_line(aes(color = year_joined.bucket), stat = 'summary', 
            fun.y = median) + 
  scale_color_brewer('Year Joined', palette = 'Set2')

```

Quiz 7 - Smoothing prop_initiated vs. tenure

# Smooth the last plot you created of
# of prop_initiated vs tenure colored by
# year_joined.bucket. You can bin together ranges
# of tenure or add a smoother to the plot.

# There won't be a solution image for this exercise.
# You will answer some questions about your plot in
# the next two exercises.

```{r}
ggplot(aes(x = tenure, y = prop_initiated), data = pf) + 
  labs(x = 'Tenure', 
       y = 'Median Proportion of Friends Initiated',
       title = 'Proportion Initiated vs. Tenure') + geom_line(aes(color = year_joined.bucket), stat = 'summary', 
            fun.y = median) + 
  geom_smooth() +
  scale_color_brewer('Year Joined', palette = 'Set2')
```
On average, the group  that initiated the greastest proportion of  Facebook friendship is that of after 2012.- see image

For the group with the largest proportion of friendships initiated, what is the group's average (mean) proportion of friendships initiated? The mean is 0.6653892
```{r}
by(pf$prop_initiated, pf$year_joined.bucket, summary)
```


Why do you think this group's proportion of friendships initiated is higher than others? They have been around for a shorter period of time. This relationship holds true for the most part across cohorts. As buckets of facebook users tenure increases, friendships initiated declines.

Quiz 10 - Price/Carat Binned, Faceted, & Colored
# Create a scatter plot of the price/carat ratio
# of diamonds. The variable x should be
# assigned to cut. The points should be colored
# by diamond color, and the plot should be
# faceted by clarity.

# The plot should look something like this.
# http://i.imgur.com/YzbWkHT.jpg.

# Note: In the link, a color palette of type
# 'div' was used to color the histogram using
# scale_color_brewer(type = 'div')

```{r}
ggplot(aes(x = cut, y = price/carat),
       data = diamonds) +
  labs(x = 'Cut', y = 'Price/Carat Ratio',
       title = 'Price/Carat Vs Cut') +
  geom_point(aes(color = color), position = 'jitter',
             alpha = 0.5) +
  scale_color_brewer(type = 'div') +
  facet_wrap( ~ clarity)
```

Quiz 11 - Gapminder Multivariate Analysis

# The Gapminder website contains over 500 data sets with information about
# the world's population. Your task is to continue the investigation you did at the
# end of Problem Set 4 or you can start fresh and choose a different
# data set from Gapminder.

# If you’re feeling adventurous or want to try some data munging see if you can
# find a data set or scrape one from the web.

# In your investigation, examine 3 or more variables and create 2-5 plots that make
# use of the techniques from Lesson 5.

# You can find a link to the Gapminder website in the Instructor Notes.

# Once you've completed your investigation, create a post in the discussions that includes:
#       1. the variable(s) you investigated, your observations, and any summary statistics
#       2. snippets of code that created the plots
#       3. links to the images of your plots


# read in the data
```{r}
fertility <- read.csv('children_per_woman_total_fertility.csv')
population <- read.csv('total_population_with_projections.csv')
co2 <- read.csv('co2_emissions_tonnes_per_person.csv')
```

### Data cleaning issues:

### Data cleaning issues:
drop tje empty colmns
1. gather the years and cases to one column each
2. remove missing incomplete cases
3. correct Column names


```{r}
install.packages(ggplot2)
install.packages("reshape")
install.packages("tidyr")
install.packages("dplyr")
library(dplyr)
library(tidyr)
library(ggplot2)

```

```{r exploring co2 data}
str(co2)

is.numeric(co2$year[0,2:216])


```


```{r}
gdURL <- "http://www.stat.ubc.ca/~jenny/notOcto/STAT545A/examples/gapminder/data/gapminderDataFiveYear.txt"
gDat <- read.delim(file = gdURL)
gDat <- droplevels(subset(gDat, continent != "Oceania"))
library(ggplot2)
library(plyr)
# http://rpubs.com/neil_spencer/stat545a-2013-hw05_spencer-nei
```


```{r}
str(gDat)
```

```{r}
summary(gDat)
```
Plotting Conditional Summaries

```{r Plotting Conditional Summaries}

ggplot(aes(x = continent, y = lifeExp),
       data = subset(gDat, !is.na(continent))) + 
  geom_boxplot() +
  stat_summary(geom = 'point', shape = 4, fun.y = mean)
ggsave('boxplot_lifeExp.png')
```

```{r}
ggplot(aes(x = continent, y = lifeExp),
       data = subset(gDat, !is.na(continent))) + 
  geom_boxplot() +
  stat_summary(geom = 'point', shape = 4, fun.y = median)
```
Write code to create a new data frame - w

```{r}

year_cont_group <- group_by(gDat, continent, year)
year_cont_group <- filter(year_cont_group, !is.na(year))
gDat_year_cont_group <- summarise(year_cont_group,
                                 mean_gdpPercap = mean(gdpPercap),
                                 median_gdpPercap = median(gdpPercap),
                                 n = n())
ungroup() %>%
arrange(gDat_year_cont_group, year)
```
```
why?

```{r}


year_country_group <- group_by(gDat, country, year)
year_country_group <- filter(year_country_group, !is.na(year))
gDat_year_country_group <- summarise(year_country_group,
                                 mean_gdpPercap = mean(gdpPercap),
                                 median_gdpPercap = median(gdpPercap),
                                 n = n())
arrange(gDat_year_country_group, year)
```











